A Simple Type Theory for Category Theory

### 2022-02-16 paper/lipics-v2021-sample-article.tex
get stuck...

\begin{document}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{lemma}{Lemma}
%% \newtheorem{construction}{Construction}
%% \newtheorem{definition}{Definition}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbbm}
\newcommand{\isaSet}{\,\,\textrm{Set}}
\newcommand{\isaTy}{\,\,\textrm{Type}}
\newcommand{\isaCat}{\,\,\textrm{Cat}}
\newcommand{\Set}{\textrm{Set}}
\newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
\newcommand{\rmod}{\text{RMod}}
\newcommand{\lmod}{\text{LMod}}
\newcommand{\varr}[2]{\text{Fun}{#1}{#2}}
\newcommand{\harr}[2]{\text{Rel}{#1}{#2}}
\newcommand{\harrapp}[3]{#1(#2;#3)}

@@

  \begin{itemize}
  \item The set $\sem{\Gamma}_{\mathcal V}$
  \item For any $\Gamma \vdash A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a set $\sem{A}_{\mathcal V} \gamma$
  \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
  \item categories...
  \item functors...
  \item profunctors...
  \item trans contexts,
  \item transformations
  \end{itemize}

  An interpretation of a signature $\Gamma$ in a virtual equipment
  $\mathcal V$, $\textrm{Interp}(\Gamma,\mathcal V)$ is an element of
  the set $\sem{\Gamma}_{\mathcal V}$ where we mutually define the set

  \begin{align*}
    \sem{\cdot}_{\mathcal V} &= 1\\
    \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
  \end{align*}

  with an 
  \begin{align*}
    \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
    \sem{\Gamma}_{\mathcal V} &= 1\\\\
    \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
    \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
  \end{align*}

@@

\subsection{Completeness}

\begin{definition}
  For any $\Gamma \isadtctx$, we construct a virtual double category
  $V^\Gamma$.
  \begin{enumerate}
  \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of ``objects'' $\Gamma
    \pipe \alpha:\cat C \vdash a : \cat D$. The category structure is
    given by substitution, with the variable $\alpha \vdash \alpha$
    the identity.
  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$
    where $d^\pm$ are the projections
  \item $V^\Gamma_h(\vec R; S)$ is the set of triples $(a,b,\Gamma
    \pipe \Phi \vdash t : S[a;b])$ where $\Phi$ is the context
    corresponding to $\vec R$ and $d^\pm$ are the obvious projections.

\subsection{Soundness}

Let $$

>>>>

\begin{document}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{lemma}{Lemma}
%% \newtheorem{construction}{Construction}
%% \newtheorem{definition}{Definition}
\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbbm}
\newcommand{\isaSet}{\,\,\textrm{Set}}
\newcommand{\isaTy}{\,\,\textrm{Type}}
\newcommand{\isaCat}{\,\,\textrm{Cat}}
\newcommand{\Set}{\textrm{Set}}
\newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
\newcommand{\rmod}{\text{RMod}}
\newcommand{\lmod}{\text{LMod}}
\newcommand{\varr}[2]{\text{Fun}{#1}{#2}}
\newcommand{\harr}[2]{\text{Rel}{#1}{#2}}
\newcommand{\harrapp}[3]{#1(#2;#3)}

@@

  \begin{itemize}
  \item The set $\sem{\Gamma}_{\mathcal V}$
  \item For any $\Gamma \vdash A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a set $\sem{A}_{\mathcal V} \gamma$
  \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
  \item For any $\Gamma \vdash \cat C \isaCat$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an object $\sem{\cat C}_{\mathcal V}\gamma$
  \item For any $\Gamma\pipe\alpha:\cat C \vdash a :\cat D$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a vertical morphism $\sem{\alpha.a}_{\mathcal V}\gamma \in \mathcal V_v(\sem{\cat C}_{\mathcal V}\gamma, \sem{\cat D}_{\mathcal V}\gamma,)$
  \item For any $\Gamma\pipe\Xi \vdash R \isaSet$ and $\gamma \in
    \sem{\Gamma}_{\mathcal V}$, a horizontal arrow $\sem{R}_{\mathcal
      V}\gamma \in \mathcal V_h$ such that $d^\pm\sem{R}_{\mathcal
      V}\gamma = \sem{d^\pm\Xi}_{\mathcal V}\gamma$. We also extend
    this in the obvious way to transformation contexts
    $\sem{\Phi}_{\mathcal V}\gamma$
  \item For any $\Gamma\pipe\Phi \vdash t : R$ and $\gamma \in
    \sem{\Gamma}_{\mathcal V}$, a 2-cell $\sem{t}_{\mathcal V}\gamma
    \in V_h(\sem{\Phi}_{\mathcal V}\gamma;\sem{R}_{\mathcal V}\gamma)$
    with $d^\pm\sem{t}_{\mathcal V}\gamma = \id$.
  \end{itemize}

  %% \begin{align*}
  %%   \sem{\cdot}_{\mathcal V} &= 1\\
  %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
  %% \end{align*}

  %% with an 
  %% \begin{align*}
  %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
  %%   \sem{\Gamma}_{\mathcal V} &= 1\\\\
  %%   \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
  %%   \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
  %% \end{align*}
\end{definition}

And we define a profunctor of interpretations of a signature in a
virtual equipment as follows:
\begin{definition}
  An interpretation of a signature $\Gamma$ in a virtual equipment
  $\mathcal V$ is an element of the set $\sem{\Gamma}_{\mathcal
    V}$. Call this set $\textrm{Interp}(\Gamma;\mathcal V)$.

  Note that this constitutes a profunctor contravariant in $\Gamma$
  and covariant in $\mathcal V$.
\end{definition}

Next, we define the syntactic virtual equipment $\textrm{Syn}(\Gamma)$
for a signature $\Gamma$.
\begin{definition}
  For any $\Gamma \isadtctx$, we construct a virtual double category
  $V^\Gamma$.
  \begin{enumerate}
  \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of objects in context
    $\Gamma \pipe \alpha:\cat C \vdash a : \cat D$. The category
    structure is given by substitution, with the variable $\alpha
    \vdash \alpha$ the identity.
  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe
    \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$ where $d^\pm$ are
    the projections
  \item $V^\Gamma_h(\vec R; S)$ is the set of triples $(a,b,\Gamma
    \pipe \Phi \vdash t : S[a;b])$ where $\Phi$ is the context
    corresponding to $\vec R$ and $d^\pm$ are the obvious projections.

Then the soundness theorem says that for any $\Gamma$, the virtual
double category $\textrm{Syn}(\Gamma)$ provides an isomorphism of profunctors:
\[ \textrm{Interp}(\Gamma, \mathcal V \cong VE(\textrm{Syn}(\Gamma), \mathcal V) \]
As we saw in a previous section, the backwards direction of this
isomorphism should be a universal interpretation:

\begin{definition}[Universal Interpretation]
  For any signature $\Gamma$ there is a ``self-interpretation'' $\eta:
  \textrm{Interp}(\Gamma, \textrm{Syn}(\Gamma))$ defined as the
  identity substitution $\id \in \sem{\Gamma}_{\textrm{Syn}(\Gamma)}$

@@
